PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2022-10-11 main.tex
Edit section 6

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/surfden_b.png}
    \caption{Feeding zone width (see equation \ref{eq:iso} required to produce the final masses for the protoplanets from the simulations listed in table \ref{tab:sim_properties}. The horizontal dashed line indicates $\tilde{b} = 2 \sqrt{3}$. Despite the vastly different initial solid surface density profiles, the feeding zone width reaches the circular orbit value around $\sim$ 60 days. \label{fig:surfden_b}}
\end{center}
\end{figure}

@@

%%%

@@

% from how far do embryos accumulate material
% do embryos migrate at all? might need to run a simulation without gas to demonstrate whether this is due to PDM
% major vs minor mergers?

@@

In all cases shown so far, the boundary between the standard
oligarchic growth and the highly-efficient short period accretion
region lies around 60 days in orbital period. The mode of accretion is
set entirely by the local value of $\alpha$, which scales with both
distance from the star and the bulk density of the planetesimals (see
equation \ref{eq:alpha}). The artificial inflation of the collision
cross section of the particles in our simulations, which is meant to
reduce the computational expense, has the side effect of reducing the
bulk densities. Because $\alpha \sim \rho_{pl}^{-1/3}$, increasing the
particle sizes by a factor of $f=6$ reduces the orbital period at which $\alpha$ has a given value by a factor of approximately 15. Therefore, one would expect the accretion boundary to lie near 5 days in orbital period for 3 g $cm^{-3}$ bodies.

Although a simulation with $f=1$ is not computationally tractable, we
can test whether the accretion boundary moves in the way we expect by
modestly changing the value of $f$. In figure \ref{fig:f6f4}, we
compare the fdHi simulation to a nearly identical run using $f=4$. In
the top panel, the embryo masses in units of the isolation mass are
shown as a function of orbital period. In the bottom panel, the value
of $\alpha$ as a function of orbital period is shown for 3 g $cm^{-3}$
bodies with an artificial radius enhancement of $f=1$, 4 and 6. The
horizontal dashed line indicates the value of alpha below which the
accretion mode switches to oligarchic growth. Comparing the top and
bottom panels, the intersection of the embryo masses and local
isolation mass matches well with the orbital period at which $\alpha
\sim 0.1$.  Also shown by the shaded region are $\alpha$ values for
realistic sized bodies with $\rho_{pl}$ between 1 and 10 g cm$^{-3}$.
% Also shown by the curved dotted lines are the $\alpha$ curves for
% realistic-sized bodies with $\rho_{pl}$ = 1, 5 and 10 g $cm^{-3}$.
In all cases, the accretion boundary still lies within the part of the disk in which solids would be expected to accumulate (citation?).

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/f6f4.png}
    \caption{In the top panel, we show the final masses of the bodies
      from the fdHi and fdHif4 simulations in units of the isolation
      mass for $\tilde{b} = 2\sqrt{3}$. The bottom panel shows the variation of $\alpha$ with orbital period for the bodies used in each case (solid curves). The orbital period at which $\alpha \simeq 0.1$ matches well with the intersection of the embryo mass distribution and isolation mass. The shaded region in the bottom panel show the values of alpha for realistic-sized ($f=1$) planetesimals with bulk densities between 10 and 1 g cm$^{-3}$.\label{fig:f6f4}}
\end{center}
\end{figure}

% compare f=4 and f=6 disk. show that the mass transition moves inwards by the correct amount as alpha changes
% how does density affect results? trappist-1 planets are 5 g/cc (agol+ 2021)

@@

In the simulations presented in this work, all collisions result in a complete merging of a pair of bodies with no loss of mass or momentum. Although simple to model, this assumption of perfect accretion may result in overly efficient growth, particularly at the innermost region of the disk where encounter velocities are the largest. To handle this issue properly, semianalytic collision resolution models have been developed and implemented in N-body codes (see \citet{leinhardt12}). However, resolving collisional debris fragments using a direct approach as we have done with {\sc ChaNGa} would result in an intractably large number of particles.

To test whether a more restrictive collision model should modify our results, we present a smaller scale test using a slightly more sophisticated collision model. In this case, a collision can result in one of two outcomes: if the impact velocity is smaller than the mutual escape velocity of the colliding particles, defined as
%TRQ: I corrected this off the top of my head: CHECK!
\begin{equation}\label{eq:v_mut}
	v_{mut, esc} = \sqrt{\frac{G (m_{1} + m_{2})}{r_{1} + r_{2}}},
\end{equation}
where $m_{1}, m_{2}$ and $r_{1}, r_{2}$ are the masses and radii of
colliding particles 1 and 2, then the bodies perfectly merge. For
impact velocities larger than $v_{mut, esc}$, no mass is transferred, and the bodies undergo a completely elastic bounce. Because there is no possibility of partial accretion, this model likely restricts growth more than the model used by \citet{leinhardt12}. However, we will show below that the bounce-merge model does not meaningfully affect the outcome of the planetesimal accretion phase, and so a partial accretion model should do the same.

For the configuration of initial conditions we have chosen, the typical encounter velocity (defined by $v_{enc} = \left< e^{2} \right>^{1/2} v_{k}$, where $v_{k}$ is the local Keplerian velocity) is about 25 percent larger than $v_{mut, esc}$. Because the encounter velocities follow a Gaussian distribution, there should be some small subset of collisions that still meet the merger criteria. In addition, $v_{mut, esc}$ becomes larger as the bodies grow and the merger criteria should become easier to meet as the system evolves.

In figure \ref{fig:frag_ecc}, we show the results of two versions of the fdLo simulation, one with mergers only (shown in orange) and one with the bounce-merge model (shown in green). The blue points in the top panel show the initial conditions used for both cases. Because the bounce-merge model greatly increases the number of timesteps required, we choose to model only the innermost part of the disk to reduce the computational cost. Although the bounce-merge simulation takes much longer to reach the same phase of evolution, the resulting orbits and masses of the embryos are indistinguishable from the merger-only case.

To investigate the differences between the two collision models early in the simulations, we show the time evolution of the ratio between the maximum and mean mass in figure \ref{fig:frag_evo}. In both cases, this ratio grows at early times, which indicates that runaway growth still operates, regardless of the collision model used. In the bounce-merge case, the mass ratio peaks at a higher value, while also undergoing a longer runaway growth phase. This suggests that the mass distribution becomes much less unimodal during this growth process, but as figure \ref{fig:frag_ecc} shows, this does not affect the resulting embryos or allow for a residual planetesimal population.

@@

% show alpha, beta study again with bounce vs merge model
% mergers still happen for 'lo' simulations but not 'hi'
% this is because the eccentricity dispersion is significantly higher in the 'hi' case
% mostly because stirring rate increases with surface density
% criteria should get less restrictive as larger bodies grow, what if we turn on bounces partway thru growth process?

\subsection{Type I Migration}

What is the migration timescale for the embryos to move to the inner edge of the disk? Maybe this section isn't terribly important and should be excluded?
TRQ: I agree

>>>>

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/surfden_b.png}
    \caption{Feeding zone width (see equation \ref{eq:iso}) required to produce the final masses for the protoplanets from the simulations listed in table \ref{tab:sim_properties}. The horizontal dashed line indicates $\tilde{b} = 2 \sqrt{3}$. Despite the vastly different initial solid surface density profiles, the feeding zone width reaches the circular orbit value around $\sim$ 60 days. \label{fig:surfden_b}}
\end{center}
\end{figure}

@@

@@

@@

In all cases shown so far, the boundary between the
oligarchic growth and the highly-efficient short period accretion
region lies around 60 days in orbital period. As discussed in section \ref{sec:sizeandhill}, the mode of accretion is
set entirely by the local value of $\alpha$, which scales with both
distance from the star and the bulk density of the planetesimals (see
equation \ref{eq:alpha}). Because we chose to artificially inflate the
collision cross section of the particles in our simulations, the bulk densities
of the particles are reduced, and the accretion boundary is shifted outward.
However,  the scaling relation between $\alpha$ and $\rho$ ($\alpha \sim \rho_{pl}^{-1/3}$) can be used
to predict where this accretion boundary should lie in a disk with realistic-sized planetesimals. The simulations
presented in this paper use a collision cross section enhancement factor of 6, which moves the boundary outward in orbital
period by a factor of approximately 15. One would therefore expect the accretion boundary to lie near 5 days in orbital period
for 3 g $cm^{-3}$ bodies.

Although a simulation with $f=1$ is not computationally tractable, we
can test whether the accretion boundary moves in the way we expect by
modestly changing the value of $f$. In figure \ref{fig:f6f4_b}, we
compare the fdHi simulation to a nearly identical run using $f=4$. In
the top panel, we show the feeding zone width required for each particle to
attain its present mass. As in figure \ref{fig:surfden_b}, we indicate the feeding
zone size predicted by oligarchic growth with a horizontal dashed line.  In the bottom panel, the value
of $\alpha$ as a function of orbital period is shown for 3 g $cm^{-3}$
bodies with an artificial radius enhancement of $f=1$, 4 and 6. The
horizontal dashed line indicates the empirical value of alpha below which the
accretion mode switches to oligarchic. Comparing the top and
bottom panels, the intersection of the feeding zone width seen in our simulations and the
feeding zone width predicted by oligarchic growth matches well with the orbital period at which $\alpha
\sim 0.1$ for both values of $f$.  Also shown by the shaded region are the expected $\alpha$ values for
realistic-sized bodies with $\rho_{pl}$ between 1 and 10 g cm$^{-3}$. Although the removal of the cross section
enhancement greatly reduces the size of the non-oligarchic region, it still should be expected to cover a portion of the
disk where planetesimals might be expected to form for a wide range of $\rho_{pl}$.

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/f6f4_b.png}
    \caption{In the top panel, we show the final masses of the bodies
      from the fdHi and fdHif4 simulations in units of the isolation
      mass for $\tilde{b} = 2\sqrt{3}$. The bottom panel shows the variation of $\alpha$ with orbital period for the bodies used in each case (solid curves). The orbital period at which $\alpha \simeq 0.1$ matches well with the intersection of the embryo mass distribution and isolation mass. The shaded region in the bottom panel show the values of alpha for realistic-sized ($f=1$) planetesimals with bulk densities between 10 and 1 g cm$^{-3}$.\label{fig:f6f4_b}}
\end{center}
\end{figure}

@@

For the simulations presented in this work, every collision results in a perfect merger between pairs of bodies, with no loss
of mass or momentum. Although simpler and less computationally expensive to model, allowing every collision to produce a
merger might result in overly efficient growth, particularly in the innermost region of the disk where the encounter velocities are largest. A proper way to handle this would be to allow for a range of collision outcomes, based on a semianalytic model (see \citet{leinhardt12}). However, resolving collisional debris, or even prolonging growth by forcing high-velocity pairs of bodies to bounce is too expensive to model, even with {\sc ChaNGa}.

To test whether a more restrictive collision model should alter the growth mode of the inner disk, we ran a smaller scale test using a slightly more restrictive collision model. In this case, a collision can result in one of two outcomes: if the impact velocity is smaller than the mutual escape velocity of the colliding particles, defined as
\begin{equation}\label{eq:v_mut}
	v_{mut, esc} = \sqrt{\frac{2 G (m_{1} + m_{2})}{r_{1} + r_{2}}},
\end{equation}
where $m_{1}, m_{2}$ and $r_{1}, r_{2}$ are the masses and radii of
colliding particles 1 and 2, then the bodies merge. For
impact velocities larger than $v_{mut, esc}$, no mass is transferred, and the bodies undergo a completely elastic bounce. Because the accretion outcome is all or nothing, this model should restrict growth more than a partial accretion model \citep{leinhardt12}. Below, we will show that the bounce-merge model does not meaningfully affect the outcome of the inner disk's planetesimal accretion phase, and so a partial accretion model should do the same.

To compare the outcome of the two collision models, we have chosen to use the initial conditions from the fdLo simulation, but 
have truncated the disk beyond 3 days in orbital period. We truncate the outer disk because the particle count with the bounce-
merge model declines much more slowly, while also requiring many more timesteps overall.
For the initial conditions we have chosen, the typical encounter velocity (defined by $v_{enc} = \left< e^{2} \right>^{1/2} v_{k}$, 
where $v_{k}$ is the local Keplerian velocity) is about 25 percent larger than $v_{mut, esc}$. Because the encounter velocities 
follow a Gaussian distribution, there should still be a small subset of collisions that still meet the merger criteria to occur early on. 
In addition, $v_{mut, esc}$ becomes larger as the bodies grow and the merger criteria should become easier to meet as the 
system evolves. For these reasons, one would expect the inhibition of growth due to the more restrictive collision model to be temporary.

In figure \ref{fig:frag_ecc}, we compare the outcomes of the simulations, one with mergers only (shown in orange) and one with the bounce-merge model (shown in green). The blue points in the top panel show the initial conditions used for both cases. Although the bounce-merge simulation takes much longer to reach the same phase of evolution, the resulting orbital properties and mass distribution of the embryos are indistinguishable from the merger-only case.

To investigate the differences in growth between the two collision models early on, we show the time evolution of the ratio between the maximum and mean mass in figure \ref{fig:frag_evo}. In both cases, this ratio first increases, which indicates that runaway growth still operates, regardless of the collision model used. In the bounce-merge case, the mass ratio peaks at a higher value, while also undergoing a longer runaway growth phase. This suggests that the mass distribution becomes much less unimodal during this growth process, but as figure \ref{fig:frag_ecc} shows, this does not affect the resulting embryos or allow for a residual planetesimal population.

@@
